---
title: "Homework 03"
author: "Sarah Tate"
format: 
  docx: 
    toc: true # includes a table of contents in rendered html format
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

## GitHub Repo Link:


## Setup for Homework:

### Loading in libraries/data from workshop

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"), # reading in the data from an excel spreadsheet
                         sheet = "First Harvest") # choosing which sheet to read within the xlsx

```

### Cleaning data frame 

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

### Correlation plots for comparing variable relationships
```{r}
ggpairs(drought_exp_clean, # data frame
        columns = c("leaf_dry_weight_g", # columns to visualize
                    "sla", 
                    "shoot_g", 
                    "root_g", 
                    "total_g"), 
        upper = list(method = "pearson")) + # calculating Pearson correlation coefficient
  theme_bw() + # cleaner theme
  theme(panel.grid = element_blank()) # getting rid of gridlines
```
### Assigning models with different variables

```{r}
# null model
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame

#diagnostics for model 0
par(mfrow = c(2, 2))
plot(model0)

# saturated model (model 1)
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

#diagnostics for model 1
par(mfrow = c(2, 2))
plot(model1)

# model 2: total mass as predicted by specific leaf area and water treatment
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

#diagnostics for model 2
par(mfrow = c(2, 2))
plot(model2)

# model 3: total mass as predicted by specific leaf area and species name
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

#diagnostics for model 3
par(mfrow = c(2, 2))
plot(model3)

# model 4: total mass as predicted by species name and water treatments
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

#diagnostics for model 4
par(mfrow = c(2, 2))
plot(model4)
```

### Model Selection to determine model of best fit
```{r}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4) # code for comparing models to each other to find model of best fit
```

### Model 4 Summary Statistics
```{r}

summary(model4)

```




## Problem 1: Multiple linear regression - model selection and construction

### a) 
  Table 1: Model Selection Data. The first column represents the model number, the second column represents which variables are present in each model, and the AIC and delta AIC values for each model are in the last two columns. Model 4 is the "best" model (AIC = -156.2, delta AIC = 0).

### b) 
  To address to central question of if total biomass (g) was influenced by specific leaf area, plant species, and/or water treatment, I assigned different models to a combination of the variables to determine which model contained the best predictor for total biomass. Model 0 included none of the variables, model 1 included all of them, and models 2-4 included combinations of 2 of the 3 variables. I used the model selection function in R to determine which model best explained the data found during experiment. It was shown that model 4 was the "best" model, which contained the variables of plant species and water treatment (AIC = -156.2, delta AIC = 0). With the largest AIC and a delta AIC of 0, it was concluded that model 4 contained the predictor variables which best explained the trends in the data. To finalize this conclusion, I determined that this model conformed to the assumptions of a linear model by creating 4 diagnostic plots to evaluate homoscedasticity, normality, and potential outliers. Through visual analysis of these plots, I determined that model 4 does conform to the assumptions of the linear model and is therefore the "best" model to predict total biomass. 

### c) 

```{r}

model_preds <- ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name")) # assigning a model for prediction data based on "best" model

model_preds_for_plotting <- model_preds %>% # new data frame for plotting prediciton data 
  rename(water_treatment = x, # renaming columns to make this easier to use
         species_name = group)


ggplot() +
  # underlying data
 geom_point(data = drought_exp_clean, # choosing data frame
             alpha = 0.1, # transparency of points
             aes(x = water_treatment,
                 y = total_g,
                color = water_treatment)) + # setting x and y variables, and choosing color for variable
  geom_point(data = model_preds_for_plotting, # choosing data frame to plot prediction data points
              aes(x = water_treatment, 
                  y = predicted,
                  color = water_treatment), # setting x and y axis variables, and color
              alpha = 1) + # more bold point for prediction data
  geom_errorbar(data = model_preds_for_plotting, # choosing data frame
            aes(x = water_treatment, 
                y = predicted, # setting x and y variables
                ymin = conf.low, 
                ymax = conf.high, # setting bounds of 95% CI
                fill = water_treatment),
            width = 0.2, # smaller width of errorbars
            alpha = 0.5) + # more transparent bars so underlying data is easier to see
  theme_classic() + # cleaner theme, no gridlines
  facet_wrap(~species_name)+  # creating different panels for species
  theme(legend.position = "none")+ # getting rid of legend
  labs(x = "Water Treatment",
       y = "Total Biomass (g)",
       title = "Figure 1: Total Biomass (g) Predicted By Plant Species And Water Treatment") + # changing x and y labels and adding title
  scale_color_manual(values = c("Well watered" = "steelblue4",
                                "Drought stressed" = "gold4")) # setting colors for each water treatment different from original ggplot colors

```


### d) 
  Figure 1: Total Biomass (g) Predicted By Plant Species and Water Treatment. The underlying data represents experimental results for total biomass of different species based on different water treatments. Bold points represent model predictions for the relationship between variables. Error bars represent a 95% confidence interval for the model predictions. (Valliere, Justin; Zhang, Jacqueline; Sharifi, M.; Rundel, Philip (2019). Data from: "Can we condition native plants to increase drought tolerance and improve restoration success?" [Dataset]. Dryad. https://doi.org/10.5061/dryad.v0861f7)

### e)


## Problem 2: Affective visualization


### a)

### b)

### c)

### d)


## Problem 3: Statistical critique


### a)

### b)

### c)

### d)





