---
title: "Homework 03"
author: "Sarah Tate"
format: 
  docx: 
    toc: true # includes a table of contents in rendered html format
execute: # for all code chunks
    warning: false # do not display warnings
    message: false # do not display messages
---

## GitHub Repo Link:


## Setup for Homework:

### Loading in libraries/data from workshop

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"), # reading in the data from an excel spreadsheet
                         sheet = "First Harvest") # choosing which sheet to read within the xlsx

```

### Cleaning data frame 

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

### Correlation plots for comparing variable relationships
```{r}
ggpairs(drought_exp_clean, # data frame
        columns = c("leaf_dry_weight_g", # columns to visualize
                    "sla", 
                    "shoot_g", 
                    "root_g", 
                    "total_g"), 
        upper = list(method = "pearson")) + # calculating Pearson correlation coefficient
  theme_bw() + # cleaner theme
  theme(panel.grid = element_blank()) # getting rid of gridlines
```
### Assigning models with different variables

```{r}
# null model
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame

#diagnostics for model 0
par(mfrow = c(2, 2))
plot(model0)

# saturated model (model 1)
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

#diagnostics for model 1
par(mfrow = c(2, 2))
plot(model1)

# model 2: total mass as predicted by specific leaf area and water treatment
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

#diagnostics for model 2
par(mfrow = c(2, 2))
plot(model2)

# model 3: total mass as predicted by specific leaf area and species name
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

#diagnostics for model 3
par(mfrow = c(2, 2))
plot(model3)

# model 4: total mass as predicted by species name and water treatments
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

#diagnostics for model 4
par(mfrow = c(2, 2))
plot(model4)
```

### Model Selection to determine model of best fit
```{r}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4) # code for comparing models to each other to find model of best fit
```


## Problem 1: Multiple linear regression - model selection and construction

### a) 
  Table 1: Model Selection Data. The first column represents the model number, the second column represents which variables are present in each model, and the AIC and delta AIC values for each model are in the last two columns. Model 4 is the "best" model (AIC = -156.2, delta AIC = 0).

### b) 
  To address to central question of if total biomass (g) was influenced by specific leaf area, plant species, and/or water treatment, I assigned different models to a combination of the variables to determine which model contained the best predictor for total biomass. Model 0 included none of the variables, model 1 included all of them, and models 2-4 included combinations of 2 of the 3 variables. I used the model selection function in R to determine which model best explained the data found during experiment. It was shown that model 4 was the "best" model, which contained the variables of plant species and water treatment (AIC = -156.2, delta AIC = 0). With the largest AIC and a delta AIC of 0, it was concluded that model 4 contained the predictor variables which best explained the trends in the data. To finalize this conclusion, I determined that this model conformed to the assumptions of a linear model by creating 4 diagnostic plots to evaluate homoscedasticity, normality, and potential outliers. Through visual analysis of these plots, I determined that model 4 does conform to the assumptions of the linear model and is therefore the "best" model to predict total biomass. 

### c) 

```{r}

model_preds <- ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name")) # assigning a model for prediction data based on "best" model

model_preds_for_plotting <- model_preds %>% # new data frame for plotting prediciton data 
  rename(water_treatment = x, # renaming columns to make this easier to use
         species_name = group)


ggplot() + # creating a plot for model 4 data
  # underlying data
 geom_point(data = drought_exp_clean, # choosing data frame
             alpha = 0.1, # transparency of points
             aes(x = water_treatment,
                 y = total_g,
                color = water_treatment)) + # setting x and y variables, and choosing color for variable
  geom_point(data = model_preds_for_plotting, # choosing data frame to plot prediction data points
              aes(x = water_treatment, 
                  y = predicted,
                  color = water_treatment), # setting x and y axis variables, and color
              alpha = 1) + # more bold point for prediction data
  geom_errorbar(data = model_preds_for_plotting, # choosing data frame
            aes(x = water_treatment, 
                y = predicted, # setting x and y variables
                ymin = conf.low, 
                ymax = conf.high, # setting bounds of 95% CI
                fill = water_treatment),
            width = 0.2, # smaller width of errorbars
            alpha = 0.5) + # more transparent bars so underlying data is easier to see
  theme_classic() + # cleaner theme, no gridlines
  facet_wrap(~species_name)+  # creating different panels for species
  theme(legend.position = "none")+ # getting rid of legend
  labs(x = "Water Treatment",
       y = "Total Biomass (g)",
       title = "Figure 1: Total Biomass (g) Predicted By Plant Species And Water Treatment") + # changing x and y labels and adding title
  scale_color_manual(values = c("Well watered" = "steelblue4",
                                "Drought stressed" = "gold4")) # setting colors for each water treatment different from original ggplot colors

```


### d) 
  Figure 1: Total Biomass (g) Predicted By Plant Species and Water Treatment. The underlying data represents experimental results for total biomass of different species based on different water treatments. Bold points represent model predictions for the relationship between variables. Error bars represent a 95% confidence interval for the model predictions. (Valliere, Justin; Zhang, Jacqueline; Sharifi, M.; Rundel, Philip (2019). Data from: "Can we condition native plants to increase drought tolerance and improve restoration success?" [Dataset]. Dryad. https://doi.org/10.5061/dryad.v0861f7)

### e) 
Plant species and water treatment best predicted total biomass (g) (Linear Model, F(7,62) = 27.08, R^2^ = 0.7257, p < 0.0001, ⍺ = 0.05). On average, with plant species held constant at the reference species Acmispon glaber (deerweed), well watered plants are larger than drought stressed plants by 0.117 ± 0.017 g. On average, plant species are 0.194 ± 0.03243 g larger than the reference species of deerweed when the water treatment is held constant. The plant species Grandelia camporum is the largest, which a total biomass which is 0.313 ± 0.03243 g larger than the references species (deerweed). 

### Model 4 Summary Statistics
```{r}
summary(model4)
```


## Problem 2: Affective visualization


### a)
  An affective visualization for my personal data would include some kind of visualization with yarn and the type of crochet projects I have been making. I have been recording the total minutes I crochet, and what kind of project I am making, as well as several other variables. I think the part of my data which is best visualized affectively is the type of project I created and how long it took me to create it. 

### b) 
sketch:

### c)
final:

### d)
  This piece shows how long each crochet project took relative to each other. Colors show whether or not the project was completed, and the height of the yarn shows how long the project took. I drew sketches of this idea on my iPad before finalizing this design using Notability on my iPad. I had to sum all of the minutes spent on each project before beginning the design process. 



## Problem 3: Statistical critique


### a)
The authors used a generalized linear mixed model to predict the probability of coral bleaching event based on a variety of environmental factors, such as latitude, depth, coral diversity, etc. Figure 2 shows the model coefficients obtained for each variable relative to their relationship to coral bleaching.
(Fig 2 image)

### b)
  The authors did a good job of visually representing their statistics, as the figure allows the reader to clearly see which predictor variables are strongly correlated with coral bleaching events. The authors included colors in the figure to show which variables has a psoitive, negative, or no contribution to coral bleaching. They used 95% confidence intervals around these predictions, but did not include the underlying data in the figure. The y axis contains all of the variable types and the x axis is the model parameter coefficients, which efficiently displays the statistical analysis. 


### c)
  While the figure is well organized, there are a lot of variables being analyzed, which makes it look intimidating to the reader. The data:ink ratio is good, with the entire graph being data and not including unnecessary analysis within the plot. Differentiating colors to provide the main results of the plot was a great choice by the author to concisely show the reader which variables had which effect on coral bleaching without much figure analysis. 


### d)
  My only complaint about this figure is the amount of variables that they chose to display on the y axis, since I think this makes the table appear more complicated than it actually is. If it is possible to combine variables, or emit the ones with no correlation to coral bleaching from the figure (but still mention them in the results or in the figure caption), I think this would make the figure more understandable at first glance.




